---
title: "Iteration & List Columns"
author: "Kimberly Lopez"
date: "2024-10-24"
output: github_document
---

```{r include=FALSE}
library(tidyverse)
library(rvest)
```

```{r}
set.seed(1)
```

# lists

```{r}
l = list(
  vec_numeric = 5:8,
  mat         = matrix(1:8, 2, 4),
  vec_logical = c(TRUE, FALSE),
  summary     = summary(rnorm(1000)))

l
```

When working with lists use square brackets to pull a vector of the list 
```{r}
l[["mat"]]
```
You can access different columns by 
```{r}
l[["mat"]][1,3]
```

And by index or position in the list 
```{r}
l[[1]]
```

# For loops

Now were going to make another list of normal samples 
```{r}
list_norms = 
  list(
    a = rnorm(20, 3, 1),
    b = rnorm(20, 0, 5),
    c = rnorm(20, 10, .2),
    d = rnorm(20, -3, 1)
  )

is.list(list_norms)
list_norms
```

If were trying to find the mean and sd of each: we can reuse a previous function we made 

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  tibble(
    mean = mean_x, 
    sd = sd_x
  )
}
```

Lets use the function to tskse the mean and sd of each sample

You can call a function based on index or by the name of the list
```{r}
mean_and_sd(list_norms[[1]])
mean_and_sd(list_norms[["a"]])
```

We can run the for loop to apply the function across multiple 

- Create output list and run a for loop
-> line 93 creates a vector type list with a length of 4 
-> we left output and the [[ ]] blank
```{r}
output = vector("list", length=4) 

for( i in 1:4){
  output[[i]]= mean_and_sd(list_norms[[i]])
}

output
```
Our output is a list of dataframes for each a,b,c,d

# `map` 

The map functions in purrr try to make the purpose of your code clear.

- functions can be passed as arguments to other functions.

We can do the same process with `map` instead

```{r}
output= map(list_norms, mean_and_sd)

output
```

We can do the same for getting median: 
```{r}
output = map_dbl(list_norms, median, .id = "input")

output
```

#`map` variants

Using map_bdl simplifies things into a vector instead of a list 
```{r}
output = map_dbl(list_norms, median, .id = "input")

output
```
```{r}
output = map_dfr(list_norms, mean_and_sd, .id = "input")

```


We can bind the rows of this list of dataframes
```{r}

output= 
  map(list_norms, mean_and_sd)|>
  bind_rows()

output
```

# List columns 

We have a datafram with columns "name" and holds the sample 
```{r}
listcol_df = 
  tibble(
    name = c("a", "b", "c", "d"),
    samp = list_norms
  )

listcol_df |> pull(name)
listcol_df |> pull(samp)
```

We can still treat this as a dataframe   
```{r}
listcol_df$samp[[1]]
```
It still treats it as a list 
we can still apply the function to the first element of the list 
```{r}
mean_and_sd(listcol_df$samp[[1]])
```
We can still map of the this list and compute mean and sd 

- map takes the input of sample from the df lists, and the function 
```{r}
map(listcol_df[["samp"]],mean_and_sd)
```

