---
title: "Iteration & List Columns"
author: "Kimberly Lopez"
date: "2024-10-24"
output: github_document
---

```{r include=FALSE}
library(tidyverse)
library(rvest)
```

```{r}
set.seed(1)
```

# lists

```{r}
l = list(
  vec_numeric = 5:8,
  mat         = matrix(1:8, 2, 4),
  vec_logical = c(TRUE, FALSE),
  summary     = summary(rnorm(1000)))

l
```

When working with lists use square brackets to pull a vector of the list 
```{r}
l[["mat"]]
```
You can access different columns by 
```{r}
l[["mat"]][1,3]
```

And by index or position in the list 
```{r}
l[[1]]
```

# For loops

Now were going to make another list of normal samples 
```{r}
list_norms = 
  list(
    a = rnorm(20, 3, 1),
    b = rnorm(20, 0, 5),
    c = rnorm(20, 10, .2),
    d = rnorm(20, -3, 1)
  )

is.list(list_norms)
list_norms
```

If were trying to find the mean and sd of each: we can reuse a previous function we made 

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  tibble(
    mean = mean_x, 
    sd = sd_x
  )
}
```

Lets use the function to tskse the mean and sd of each sample

You can call a function based on index or by the name of the list
```{r}
mean_and_sd(list_norms[[1]])
mean_and_sd(list_norms[["a"]])
```

