---
title: "Iteration & List Columns"
author: "Kimberly Lopez"
date: "2024-10-24"
output: github_document
---

```{r include=FALSE}
library(tidyverse)
library(rvest)
```

```{r}
set.seed(1)
```

# lists

```{r}
l = list(
  vec_numeric = 5:8,
  mat         = matrix(1:8, 2, 4),
  vec_logical = c(TRUE, FALSE),
  summary     = summary(rnorm(1000)))

l
```

When working with lists use square brackets to pull a vector of the list 

```{r}
l[["mat"]]
```

You can access different columns by 

```{r}
l[["mat"]][1,3]
```

And by index or position in the list 
```{r}
l[[1]]
```

# For loops

Now were going to make another list of normal samples 

```{r}
list_norms = 
  list(
    a = rnorm(20, 3, 1),
    b = rnorm(20, 0, 5),
    c = rnorm(20, 10, .2),
    d = rnorm(20, -3, 1)
  )

is.list(list_norms)
list_norms
```

If were trying to find the mean and sd of each: we can reuse a previous function we made 

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  tibble(
    mean = mean_x, 
    sd = sd_x
  )
}
```

Lets use the function to tskse the mean and sd of each sample

You can call a function based on index or by the name of the list

```{r}
mean_and_sd(list_norms[[1]])
mean_and_sd(list_norms[["a"]])
```

We can run the for loop to apply the function across multiple 

- Create output list and run a for loop
- line 93 creates a vector type list with a length of 4 
- we left output and the [[ ]] blank

```{r}
output = vector("list", length=4) 

for( i in 1:4){
  output[[i]]= mean_and_sd(list_norms[[i]])
}

output
```

Our output is a list of dataframes for each a,b,c,d

# `map` 

The map functions in purrr try to make the purpose of your code clear.

- functions can be passed as arguments to other functions.

We can do the same process with `map` instead

```{r}
output= map(list_norms, mean_and_sd)

output
```

We can do the same for getting median: 

```{r}
output = map_dbl(list_norms, median, .id = "input")

output
```

#`map` variants

Using map_bdl simplifies things into a vector instead of a list 

```{r}
output = map_dbl(list_norms, median, .id = "input")

output
```
```{r}
output = map_dfr(list_norms, mean_and_sd, .id = "input")

```


We can bind the rows of this list of dataframes

```{r}

output= 
  map(list_norms, mean_and_sd)|>
  bind_rows()

output
```

# List columns 

We have a datafram with columns "name" and holds the sample 

```{r}
listcol_df = 
  tibble(
    name = c("a", "b", "c", "d"),
    samp = list_norms
  )

listcol_df |> pull(name)
listcol_df |> pull(samp)
```

We can still treat this as a dataframe  

```{r}
listcol_df$samp[[1]]
```
It still treats it as a list 
we can still apply the function to the first element of the list

```{r}
mean_and_sd(listcol_df$samp[[1]])
```
We can still map of the this list and compute mean and sd 

- map takes the input of sample from the df lists, and the function 

```{r}
map(listcol_df[["samp"]],mean_and_sd)
```

We can use the map function to **add a list column** 

- map_dbl will give you the number
- we can use `unnest(output)` to get rid of a nested sample but keeping the values of the tibble from `mean_and_sd`

```{r}
listcol_df= 
  listcol_df|>
  mutate(
    output=map(samp,mean_and_sd),
    iqr=map_dbl(samp,IQR))|>
  unnest(output)

listcol_df
```

# NSDUH 

We want to use a version of our function we saved in our source document last time:

```{r}
nsduh_table_format = function  (html, table_num) {
  
  out_table = 
    html|>
    html_table()|>
    nth(table_num)|>
    slice(-1)|>
    select(-contains("P Value"))
  
  return(out_table)
}
```


We need to import the html, and then extract the correct tables

```{r}

nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

nsduh_table_format(html= nsduh_html,table_num = 1)
nsduh_table_format(html= nsduh_html,table_num = 4)
nsduh_table_format(html= nsduh_html,table_num = 5)
```

Creating a column and putting it into a dataframe

```{r}
nsduh_df= 
  tibble(
    drug= c("marj","cocaine","herion"),
    table_n= c(1,4,5)
  )|>
  mutate(table=map(table_n,nsduh_table_format, html = nsduh_html))|>
  unnest(table)

head(nsduh_df)
```

We can either use a for loop or use the map function to create another df 

**For loop way**

```{r}
output = vector("list", 3)

for (i in c(1, 4, 5)) {
  output[[i]] = nsduh_table_format(nsduh_html, i)
}

nsduh_results = bind_rows(output)
head(nsduh_results,n=5)
```

**`map` way **

```{r}
nsduh_results = 
  map(c(1, 4, 5), nsduh_table_format, html = nsduh_html) |> 
  bind_rows()

head(nsduh_results,n=5)
```

# Operations on nested data 

Were importing 3 weather stations and creating a df of it 
```{r message=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Create a list column to nest everything from date to tmin into a column

```{r}
weather_nest= 
  weather_df|>
  nest(data= date:tmin)
weather_nest
```

This is a list column so we can extract the data column which will return 3 seperate dataframes for each station: 
```{r}
weather_nest[["data"]]
```


OR we can select dataframe of only from 1 
```{r}
weather_nest[["data"]][[1]]
```

We can try a regression tmax on tmin for 3 stations by doing a forloop or 

**for loop method**

We created a new column using `mutate()` and then used the `\(x)` anonymous feature since `lm` is more complicated and requires a data specification for each lm 

- `pull()` allows you to pull all 3 model fits

```{r}

weather_nest|>
  mutate(
    model_fit=map(data,\(x) lm(tmax ~ tmin,data=x))
  )|>
  pull(model_fit)
```

** This can also be done by ** 

```{r}
weather_lm(weather_nest[["data"]][[1]])

weather_nest|>
  mutate(model_fit = map(data,weather_lm))
```

