---
title: "simulation"
author: "Kimberly Lopez"
date: "2024-10-31"
output: github_document
---

**Example**

```{r}
library(tidyverse)

set.seed(1)

```
## Simulation: Mean and SD for one ð‘›


```{r}
sim_df = 
  tibble(
    x=rnorm(30,10,5)
  )

sim_df|>
  summarize(
    mean= mean(x),
    sd= sd(x)
  )
```
We will be using the function; 
```{r}
sim_mean_sd = function(n, mu = 2, sigma = 3) {
  
  sim_data = 
    tibble(
    x = rnorm(n, mean = mu, sd = sigma),
  )
  
  sim_data |> 
    summarize(
      samp_mean = mean(x),
      samp_sd = sd(x)
    )
}
```

Lets run a simulation 100 times to see if this will actually be a normal distribution 

- since we had a list of dataframes, we could use Bind() to combine them all with each row being an entry 
- we can also summarize on these inputs to take the average mean and sd for all 100 samples
```{r}
output = vector("list", 100)

for (i in 1:100) {
  output[[i]] = sim_mean_sd(30)
}

sim_results = bind_rows(output)|>
  summarize(
    av_samp_mean = mean(samp_mean),
    SE_samp_mean = sd(samp_mean)
  )
sim_results
```
**we can do the same thing using mapping instead of a forloop to do this** 

- first create a df to keep all the data 
- map over an iteration
- we use mutate to update our empty tibble with these results 
```{r}
sim_results= 
  tibble(
    iter = 1:100
  )|>
  mutate(samp_results = map(iter , sim_mean_sd, n= 30))|>
  unnest(samp_results)

sim_results
```

**We can try different sample size using `expand_grid()**

- the list of inputs into our function they will iterate through the different sample sizes n 
```{r}
sim_results= 
  expand_grid(
    sample_size= c(10,30,60,100),
    iter= 1:100
  )|>
  mutate(samp_results= map(sample_size,sim_mean_sd))|>
  unnest(samp_results)

sim_results
```

Now we can  `group_by` n and summarize the stanadrd error of the sample mean 

```{r}
sim_results |> 
  group_by(sample_size)|>
  summarize(
    se= sd(samp_mean)
  )
```


**we can also visualize these results**

-> used a histogram since a `violin` requires you to have a factor variable
```{r}
sim_results|> 
  filter(sample_size==100)|>
  ggplot(aes(x=samp_mean))+
  geom_histogram()
```
# Varying two simulation paramaters
```{r echo=TRUE}
sim_results|> 
  mutate(
    sample_size = str_c("n = ", sample_size),
    sample_size = fct_inorder(sample_size)) |> 
  ggplot(aes(x = sample_size, y = samp_mean, fill= sample_size)) +
  geom_violin()
```


# Simple Linear Regression 

```{r}
 sim_data= 
  tibble(
    x= rnorm(30, mean = 1, sd= 1),
    y = 2+3*x+rnorm(30,0,1)
  )

sim_data|>
  ggplot(aes(x=x, y=y))+
  geom_point()+
  stat_smooth(method="lm")

lm_fit= lm(y~x, data= sim_data)
```


## Simulation: Simple Linear Regression for one n 
**We can turn this into a function that returns the coefficients output**

- `coef(lm_fit[index#])` will give you the beta coefficients

```{r}
sim_regression = function (n){
  
  sim_data= 
  tibble(
    x= rnorm(n, mean = 1, sd= 1),
    y = 2+3*x+rnorm(n,0,1)
  )
  
  lm_fit = lm(y~x, data = sim_data)
  
  out_df = 
    tibble( 
      beta0_hat = coef(lm_fit)[1],
      beta1_hat = coef(lm_fit)[2])
  return(out_df)
}
```

Now we can run this function over many 

```{r}
sim_results = 
  expand_grid(
    sample_size = 30 , 
    iter= 1:100
  )|>
  mutate(lm_res = map(sample_size, sim_regression))|>
  unnest(lm_res)

head(sim_results)

sim_results|>
  mutate(sample_sez= str_c("n = ", sample_size)) |>
  ggplot(aes(x= sample_size, y = beta1_hat))+ 
  geom_boxplot()

sim_results |> 
  ggplot(aes(x = beta0_hat, y = beta1_hat)) + 
  geom_point()
```

## Birthday Problem 

```{r}

bday_sim = function(n){
  bdays= sample( 1:365, size = 10, replace= TRUE)
  
  duplicate = length(unique(bdays))<10
  
  return(duplicate)
}

bday_sim(10)
```

`map_lgl` gives you the logical return 

- we can use code to get the average of true or false based on sample size
- exapand grid helps us iterate the map function to simulate 
```{r}
sim_results = 
  expand_grid(
    n= c(10,50),
    iter= 1:100
  )|>
  mutate(res = map_lgl(n, bday_sim))|>
  group_by(n)|>
  summarize(prob = mean(res))

sim_results
```

# Varying Two simulation paramaters 


