---
title: "Writing Functions"
author: "Kimberly Lopez"
date: "2024-10-24"
output: github_document
---

If you use the same code twice, you need a function – this will improve code readability, facilitate troubleshooting, and reduce chances for mistakes. This content looks at the best approaches for writing R functions.

# Example 

**setting the seed so that the output on this page is fixed**

```{r}
library(tidyverse)
library(rvest)
set.seed(1)

```

## My First Function 

Generating random numbers from the normal distribution and finidng the standard deviation. 
```{r}
x_vec = rnorm(25, mean = 5, sd = 3)

(x_vec - mean(x_vec)) / sd(x_vec)
```
To create a function to do this we create a new function that takes the sample as an argument, computes the vector of Z scores in the body, and returns the result. 

- x is the generic input of the sample 
- can be explicit and state to return(z)

```{r}
z_scores = function(x) {
  
  z = (x - mean(x)) / sd(x)
  z
  
}

z_scores(x=x_vec)
```
Checking to see if it works: 
```{r}
z_scores(x=3)
```

To update the code so it works always: 

- set a condition if x is numeric, and a minimum of 1 vector to compute the SD
```{r}
z_scores = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Z scores cannot be computed for length 1 vectors")
  }
  
  z = (x- mean(x)) / sd(x)
  
  return(z)
}

```

## Multiple Outputs

In some cases it might be better to return the mean and standard deviation instead of the Z scores. A first option is to store each of the values in a named list, and to return that list. 

- creating a new mean and standard deviation
- saves mean and sd in a tibble dataframe and returns if
```{r}
mean_and_sd= function(x){
  mean_x=mean(x)
  sd_x=sd(x)
  out_df = tibble(mean= mean_x,
                  sd= sd_x)
  return(out_df)
}
```

Apply to our vector x_vec
```{r}
mean_and_sd(x_vec)
```
## Checking code using simulation 

```{r}
sim_df = 
  tibble(
    x=rnorm(30,10,5)
    )

sim_df|>
  summarize(
    mean= mean(x),
    sd= sd(x)
  )
```
** **Creating a simulation function to check sample mean and sd:** **

We want a function to take inputs: 

- sample size n 
- mean mu
- standard deviation sigma 
- check code by running n = 30, mu= 4, sigma=12
```{r}
sim_mean_sd = function(n, mu, sigma ) {
  
  sim_df = tibble(
    x = rnorm(n, mean = mu, sd = sigma),
  )
  
  out_df=
    sim_df |> 
    summarize(
      mu_hat = mean(x),
      sigma_hat = sd(x)
    )
  
  return(out_df)
}

sim_mean_sd( n=30,mu=4, sigma=12)
```

