---
title: "Writing Functions"
author: "Kimberly Lopez"
date: "2024-10-24"
output: github_document
---

If you use the same code twice, you need a function – this will improve code readability, facilitate troubleshooting, and reduce chances for mistakes. This content looks at the best approaches for writing R functions.

# Example 

**setting the seed so that the output on this page is fixed**

```{r}
library(tidyverse)
library(rvest)
set.seed(1)

```

## My First Function 

Generating random numbers from the normal distribution and finidng the standard deviation. 
```{r}
x_vec = rnorm(25, mean = 5, sd = 3)

(x_vec - mean(x_vec)) / sd(x_vec)
```
To create a function to do this we create a new function that takes the sample as an argument, computes the vector of Z scores in the body, and returns the result. 

- x is the generic input of the sample 
- can be explicit and state to return(z)

```{r}
z_scores = function(x) {
  
  z = (x - mean(x)) / sd(x)
  z
  
}

z_scores(x=x_vec)
```
Checking to see if it works: 
```{r}
z_scores(x=3)
```

To update the code so it works always: 

- set a condition if x is numeric, and a minimum of 1 vector to compute the SD
```{r}
z_scores = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Z scores cannot be computed for length 1 vectors")
  }
  
  z = (x- mean(x)) / sd(x)
  
  return(z)
}

```

## Multiple Outputs

In some cases it might be better to return the mean and standard deviation instead of the Z scores. A first option is to store each of the values in a named list, and to return that list. 

- creating a new mean and standard deviation
- saves mean and sd in a tibble dataframe and returns if
```{r}
mean_and_sd= function(x){
  mean_x=mean(x)
  sd_x=sd(x)
  out_df = tibble(mean= mean_x,
                  sd= sd_x)
  return(out_df)
}
```

Apply to our vector x_vec
```{r}
mean_and_sd(x_vec)
```
## Checking code using simulation 

```{r}
sim_df = 
  tibble(
    x=rnorm(30,10,5)
    )

sim_df|>
  summarize(
    mean= mean(x),
    sd= sd(x)
  )
```
** **Creating a simulation function to check sample mean and sd:** **

We want a function to take inputs: 

- sample size n 
- mean mu
- standard deviation sigma 
- check code by running n = 30, mu= 4, sigma=12

**You can set paramaters equal to default values by setting function (n=30,mu=2,sigma=2) but you can always override it by putting different values in when you run the code on inputs**
```{r}
sim_mean_sd = function(n, mu, sigma ) {
  
  sim_df = tibble(
    x = rnorm(n, mean = mu, sd = sigma),
  )
  
  out_df=
    sim_df |> 
    summarize(
      mu_hat = mean(x),
      sigma_hat = sd(x)
    )
  
  return(out_df)
}

sim_mean_sd( n=30,mu=4, sigma=12)
```


# Revisiting past examples: LOTR DATA

```{r}
fellowship_ring = readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") |>
  mutate(movie = "fellowship_ring")

two_towers = readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") |>
  mutate(movie = "two_towers")

return_king = readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") |>
  mutate(movie = "return_king")

lotr_tidy = bind_rows(fellowship_ring, two_towers, return_king) |>
  janitor::clean_names() |>
  pivot_longer(
    female:male,
    names_to = "sex",
    values_to = "words") |> 
  mutate(race = str_to_lower(race)) |> 
  select(movie, everything()) 
```

**Write a function that can be used to abstract the data loading and cleaning process. Use this function to recreate the tidied LoTR dataset**

--> if we start from loading dataset, the input values have to be path, range, and movie_name

--> body of the function should tidy the data clean, pivot, mutate

```{r}

lotr_load_and_tidy = function(path, range, movie_name) {
  
  movie_df = 
    readxl::read_excel(path, range = range) |>
    janitor::clean_names() |>
    pivot_longer(
      female:male,
      names_to = "sex",
      values_to = "words") |>
    mutate(
      race = str_to_lower(race),
      movie = movie_name) |> 
    select(movie, everything())
  
  return(movie_df)
  
}
```

To check apply the function to create **lotr_tidy** dataframe by binding the output of the functions

```{r}
lotr_tidy = 
  bind_rows(
    lotr_load_and_tidy("data/LotR_Words.xlsx", "B3:D6", "fellowship_ring"),
    lotr_load_and_tidy("data/LotR_Words.xlsx", "F3:H6", "two_towers"),
    lotr_load_and_tidy("data/LotR_Words.xlsx", "J3:L6", "return_king"))

head(lotr_tidy)
```

## # Revisiting past examples: NSDUH Data 

This is a reading data from the web code that scrapes the information for us. 

We could take the first table "nth(1)" by: **read_html(html)**
```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

data_marj = 
  nsduh_html |> 
  html_table() |> 
  nth(1) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```

We can also create a function that does the data import for us: 

--> takes the input html for the link 
--> take the input table_num that we put into "nth()" to ddenote which table from the link we want 
--> takes the input "table_name" to know which 
```{r}
nsduh_table <- function(html, table_num, table_name) {
  
  table = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent),
      name = table_name) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  table
  
}
```


```{r}
nsduh_results = 
  bind_rows(
    nsduh_table(nsduh_html, 1, "marj_one_year"),
    nsduh_table(nsduh_html, 4, "cocaine_one_year"),
    nsduh_table(nsduh_html, 5, "heroin_one_year")
  )

head(nsduh_results)
```

**We do not want to put the html link within our function because it will constantly download the data from the internet multiple times, so name this outside the functions**


We can save this nsduh function into a r script paste it there save that into a "source folder" and then call on the function whenevre in this project and it does the same thing.

```{r}
source("source/nsduh_table_format.R")

bind_rows(
    nsduh_table(nsduh_html, 1, "marj_one_year"),
    nsduh_table(nsduh_html, 4, "cocaine_one_year"),
    nsduh_table(nsduh_html, 5, "heroin_one_year")
  )
```

